Variational Monte Carlo on Bosonic Systems

### 2018-04-06 project_1/article/article.tex
Spell check section on initial distribution

    \subsection{One-Body Densities}
        In order to make some conclusion towards where the particles in our
        systems are most prone to exist, we need a good way to visualise our
        wavefunctions. If we are to visualize the combined wavefunction of two
        three-dimensional particles we need four dimensions per particle and can
        subtract one dimension as we look at a single probability density
        yielding a grand total of seven dimensions. The fact that we will be
        studying systems of up to $N = 500$ particles only adds to the headache.
        We therefore introduce the single-particle density function, or
        \emph{one-body density} $\rho(\vf{r})$. The one-body density is defined
        by the integral over all coordinates except one,
        \begin{align}
            \rho(\vf{r}_1)
            \equiv
            \int \dd\vf{r}_2 \dots \dd\vf{r}_N
            \abs{\Psi(\vf{r}_1, \dots, \vf{r}_N)}^2.
        \end{align}
        The one-body density represents the distribution of all the particles in
        the system. Furthermore, $\rho(\vf{r}_1)\dd\vf{r}_1$ will yield the
        probability of finding any particle within the volume element
        $\dd\vf{r}_1$. By convention the one-body density should be normalized
        such that the following relation holds,
        \begin{align}
            N = \int \rho(\vf{r}_1)\dd\vf{r}_1.
        \end{align}
        In other words, the integral over the entire volume in
        $\vf{r}_1$-direction gives us the total number of particles in our
        system. This can be a bit odd at first sight as we are used to getting
        unity when we integrate over the entire probability
        density.\cite{hogberget2013quantum}

@@

        \subsubsection{Initial distribution of particles}
            Prior to performing the Metropolis-Hastings algorithm we have to an
            initial distribution of the particles. As the algorithm is supposed
            to move particles in the right direction this distribution can be
            \emph{somewhat} arbitrary. There are however a few caveats to this
            approach.

            \begin{enumerate}
                \item The distribution is unphysical. This results in the
                    particles taking up positions which can yield bogus values
                    for the local energy. A remedy is to allow for some
                    \emph{thermalization steps}, i.e., scatter the particles and
                    run the algorithm for a set number of steps prior to
                    starting the actual sampling.
                \item The initial \emph{spread} of the particles is too small or
                    too large. The former can lead to an acceptance ratio of
                    zero if no steps are accepted due to overlap. This is
                    something that can happen to the interacting system as an
                    overlap between the particles makes the wavefunction
                    evaluate to zero thus rejecting all steps in the ratio test.
                    The latter can lead to the system never reaching a true
                    state as they are scattered too far away from the
                    interesting regions in the configuration space.
                \item The system imposes some restrictions on the distribution.
                    This can mean that we have to distribute the particles in a
                    certain way to make sure that we don't break the
                    restrictions. This resembles the former case when the spread
                    is too small for our systems.
            \end{enumerate}

>>>>

    \subsection{One-Body Densities}
        In order to make some conclusion towards where the particles in our
        systems are most prone to exist, we need a good way to visualise our
        wavefunctions. If we are to visualize the combined wavefunction of two
        three-dimensional particles we need four dimensions per particle and can
        subtract one dimension as we look at a single probability density
        yielding a grand total of seven dimensions. The fact that we will be
        studying systems of up to $N = 500$ particles only adds to the headache.
        We therefore introduce the single-particle density function, or
        \emph{one-body density} $\rho(\vf{r})$. The one-body density is defined
        by the integral over all coordinates except
        one\cite{kristiansen2017time},
        \begin{align}
            \rho(\vf{r}_1)
            \equiv
            \int \dd\vf{r}_2 \dots \dd\vf{r}_N
            \abs{\Psi(\vf{r}_1, \dots, \vf{r}_N)}^2.
        \end{align}
        The one-body density represents the distribution of all the particles in
        the system. Furthermore, $\rho(\vf{r}_1)\dd\vf{r}_1$ will yield the
        probability of finding any particle within the volume element
        $\dd\vf{r}_1$. By convention the one-body density should be normalized
        such that the following relation holds,
        \begin{align}
            N = \int \rho(\vf{r}_1)\dd\vf{r}_1.
        \end{align}
        In other words, the integral over the entire volume in
        $\vf{r}_1$-direction gives us the total number of particles in our
        system. This can be a bit odd at first sight as we are used to getting
        unity when we integrate over the entire probability
        density.\cite{hogberget2013quantum}

@@

        \subsubsection{Initial distribution of particles}
            Prior to performing the Metropolis-Hastings algorithm we have to
            have an initial distribution of the particles. As the algorithm is
            supposed to move particles in the right direction this distribution
            can be \emph{somewhat} arbitrary. There are however a few caveats to
            this approach.

            \begin{enumerate}
                \item The distribution is unphysical. This results in the
                    particles taking up positions which can yield bogus values
                    for the local energy. A remedy is to allow for some
                    \emph{thermalization steps}, i.e., scatter the particles and
                    run the algorithm for a set number of steps prior to
                    starting the actual sampling.

                \item The initial \emph{spread} of the particles is too small or
                    too large. The former can lead to an acceptance ratio of
                    zero if no steps are accepted due to overlap. This is
                    something that can happen to the interacting system we are
                    studying as an overlap between the particles makes the
                    wavefunction evaluate to zero thus rejecting all steps in
                    the ratio test.  The latter can lead to the system never
                    reaching a true state as they are scattered too far away
                    from the interesting regions in the configuration space.

                \item The system imposes some restrictions on the distribution.
                    This can mean that we have to distribute the particles in a
                    certain way to make sure that we don't break the
                    restrictions. This resembles the former case when the spread
                    is too small for our systems.
            \end{enumerate}
